---
title: "Real Estate Properties Modeling Project Slides"
subtitle: <small>"That's what I love about these GAMs, man. My RMSE keeps getting lower and the coverage stays the same." --JB<small>
author: "Beau Coker, Muyao Sun, Jiahui Wang, Huijia Yu"
date: "4/30/2017"
output: beamer_presentation
fontsize: 5pt
---

```{r, include=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(mgcv))
suppressMessages(library(ggplot2))
suppressMessages(library(reshape2))
suppressMessages(library(glmnet))
suppressMessages(library(gridExtra))
suppressMessages(library(randomForest))
```

```{r, include=FALSE}
clean_df <- function(df, na_to_none, na_to_zero) {
  
  # Change to log price
  df$price <- log(df$price)
  
  # Garage Year Built
  df <- df %>%
    mutate(Garage.Era = as.factor(cut(Garage.Yr.Blt, breaks=5, labels=FALSE))) %>%
    select(-Garage.Yr.Blt)
  
  
  # For features with absense of level (e.g., no garage)
  # replace NA with "NONE"
  for (i in 1:length(na_to_none)) {
    levels_before <- levels(df[[na_to_none[i]]])
    df[[na_to_none[i]]] <- addNA(df[[na_to_none[i]]])
    levels(df[[na_to_none[i]]]) <- c(levels_before, "None")
  }
  
  # NAs to 0
  for (i in 1:length(na_to_zero)) {
    id_na <- is.na(df[[na_to_zero[i]]])
    df[[na_to_zero[i]]][id_na] <- 0
  }
  
  # Fix specific ""
  df$Mas.Vnr.Type[df$Mas.Vnr.Type==""] <- "None"
  df$Mas.Vnr.Type <- factor(df$Mas.Vnr.Type)
  
  df$Electrical[df$Electrical==""] <- "SBrkr"
  df$Electrical <- factor(df$Electrical)
  
  df$Bsmt.Exposure[df$Bsmt.Exposure==""] <- "No"
  df$Bsmt.Exposure <- factor(df$Bsmt.Exposure)
  
  df$BsmtFin.Type.2[df$BsmtFin.Type.2==""] <- "Unf"
  df$BsmtFin.Type.2 <- factor(df$BsmtFin.Type.2)
  
  df$Garage.Finish[df$Garage.Finish==""] <- "Unf"
  df$Garage.Finish <- factor(df$Garage.Finish)
  
  ### Add/Combine new columns
  df <- df %>%
    mutate(N.Floors = ifelse(X2nd.Flr.SF>0,2,1)) %>%
    mutate(Total.SF = X1st.Flr.SF + X2nd.Flr.SF + Total.Bsmt.SF) %>%
    mutate(Total.Bsmt.SF = ifelse(Total.Bsmt.SF>0, TRUE, FALSE)) %>%
    mutate(Bsmt = ifelse(Total.Bsmt.SF>0,TRUE,FALSE)) %>% 
    mutate(Low.Qual.Fin = ifelse(Low.Qual.Fin.SF>0,TRUE,FALSE)) %>%
    mutate(Porch.SF = Wood.Deck.SF + Open.Porch.SF + Enclosed.Porch + 
             X3Ssn.Porch + Screen.Porch) %>%
    mutate(Porch = ifelse(Porch.SF>0,TRUE,FALSE)) %>%
    mutate(Pool = ifelse(Pool.Area>0,TRUE,FALSE)) %>%
    mutate(Full.Bath = Full.Bath + Bsmt.Full.Bath) %>%
    mutate(Half.Bath = Half.Bath + Bsmt.Half.Bath)
  
  ### Remove columns
  df <- df %>%
    select(-area, -X1st.Flr.SF, -X2nd.Flr.SF, -Total.Bsmt.SF, -TotalSq, -Low.Qual.Fin.SF,
           -Wood.Deck.SF,-Open.Porch.SF,-Enclosed.Porch,-X3Ssn.Porch,-Screen.Porch,
           -Bsmt.Full.Bath, -Bsmt.Half.Bath) 
  
  
  # Remove all unused "" levels
  id.fac <- which(sapply(df,is.factor))
  for (i in id.fac) {
    df[[i]] <- factor(df[[i]])
  }
  
  return(df)
}
```


```{r, include=FALSE}
load("ames_train.Rdata")
load("ames_test.Rdata")
load("ames_validation.Rdata")

# For features with absense of level (e.g., no garage)
# replace NA with "NONE"
na_to_none <- c("Alley",
                "Bsmt.Qual",
                "Bsmt.Exposure",
                "Bsmt.Cond",
                "Bsmt.Exposure",
                "BsmtFin.Type.1",
                "BsmtFin.Type.2",
                "Fireplace.Qu",
                "Garage.Type",
                "Garage.Finish",  
                "Garage.Qual",     
                "Garage.Cond",
                "Garage.Era", # New variable
                "Pool.QC",
                "Fence",
                "Misc.Feature")

na_to_zero <- c("Lot.Frontage",
                "Mas.Vnr.Area",
                "Bsmt.Full.Bath",
                "Bsmt.Half.Bath")

ames_all <- clean_df(
  rbind(
    mutate(ames_train,dataset="train"),
    mutate(ames_test,dataset="test"),
    mutate(ames_validation,dataset="validation")
  ),na_to_none,na_to_zero
)

ames_train <- ames_all %>% filter(dataset=="train") %>% select(-dataset)
ames_test <- ames_all %>% filter(dataset=="test") %>% select(-dataset)
ames_validation <- ames_all %>% filter(dataset=="validation") %>% select(-dataset, -price)
```

```{r, include=FALSE}
### Ordinal Factors to Numeric


ord_to_num_Ex_Po = function(x){
  # For factors with "Ex"   "Fa"   "Gd"   "Po"   "TA"   
  levels(x) <- c("Ex"=5, "Fa"=3, "Gd"=4, "Po"=1, "TA"=2)
  return (as.numeric(as.character(x)))
}

ord_to_num_Ex_NA = function(x) {
  # For factors with "Ex"   "Fa"   "Gd"   "Po"   "TA"   "None"
  levels(x) <- c("Ex"=6, "Fa"=4, "Gd"=5, "Po"=2, "TA"=3, "None"=1)
  return (as.numeric(as.character(x)))
}


#turning ordinal variables into numeric

ordinal_to_num = function(df, Ex_Po, Ex_NA){
  ### Generic ones first 
  for (i in 1:length(Ex_Po)) {
    df[[Ex_Po[i]]] <- ord_to_num_Ex_Po(df[[Ex_Po[i]]])
  }
  for (i in 1:length(Ex_NA)) {
    df[[Ex_NA[i]]] <- ord_to_num_Ex_NA(df[[Ex_NA[i]]])
  }
  

  ### Specific ones
  levels(df$Bsmt.Exposure) <- c("Av"=4, "Gd"=5, "Mn"=3, "No"=2, "None"=1)
  df$Bsmt.Exposure<- as.numeric(as.character(df$Bsmt.Exposure))
  
  levels(df$BsmtFin.Type.1) <- c("ALQ"=6, "BLQ"=5, "GLQ"=7, "LwQ"=3, "Rec"=4, "Unf"=2, "None"=1)
  df$BsmtFin.Type.1<- as.numeric(as.character(df$BsmtFin.Type.1))
  
  levels(df$BsmtFin.Type.2) <- c("ALQ"=6, "BLQ"=5, "GLQ"=7, "LwQ"=3, "Rec"=4, "Unf"=2, "None"=1)
  df$BsmtFin.Type.2<- as.numeric(as.character(df$BsmtFin.Type.2))
  
  levels(df$Garage.Finish) <- c("Fin"=4,  "RFn"=3,  "Unf"=2,  "None"=1)
  df$Garage.Finish<- as.numeric(as.character(df$Garage.Finish))
  
  levels(df$Garage.Finish) <- c("Ex"=5, "Fa"=3, "Gd"=4, "TA"=2, "None"=1)
  df$Garage.Finish<- as.numeric(as.character(df$Garage.Finish))
  
  
  levels(df$Lot.Shape)<- c(3, 2, 1, 4)
  df$Lot.Shape<- as.numeric(as.character(df$Lot.Shape))
  
  #also missing ELO factor
  levels(df$Utilities)<- c(4, 2, 3, 1)
  df$Utilities<- as.numeric(as.character(df$Utilities))
  
  levels(df$Land.Slope)<- c(3, 2, 1)
  df$Land.Slope<- as.numeric(as.character(df$Land.Slope))
  
  #because no "Po" values in the training data
  levels(df$Exter.Qual)<- c(5, 2, 4, 3, 1)
  df$Exter.Qual<- as.numeric(as.character(df$Exter.Qual))
  
  #levels(df$Functional)<- c(4, 3, 7, 6, 5, 1, 2, 8)
  #df$Functional<- as.numeric(as.character(df$Functional))
  
  df$Paved.Drive<- as.numeric(df$Paved.Drive)
  return (df)
}

Ex_Po <- c("Exter.Cond","Heating.QC", "Kitchen.Qual")
Ex_NA <- c("Bsmt.Qual","Bsmt.Cond","Fireplace.Qu", "Garage.Qual", "Garage.Cond")

ames_train<- ordinal_to_num(ames_train, Ex_Po, Ex_NA)
ames_test<- ordinal_to_num(ames_test, Ex_Po, Ex_NA)
ames_validation<- ordinal_to_num(ames_validation, Ex_Po, Ex_NA)
```


## Location, location, location {.smaller}

```{r, echo=FALSE}
### Neighborhood boxplots

df_plot <- ames_train %>% select(price, Neighborhood)

# Order neighborhoods by mean price
lvl.ordered <- df_plot %>%
  group_by(Neighborhood) %>%
  dplyr::summarize(mean_price = mean(price)) %>%
  arrange(mean_price) %>%
  .[[1]]

df_plot$Neighborhood <- factor(df_plot$Neighborhood, levels=lvl.ordered)

ggplot(df_plot, aes(y=price, x=Neighborhood)) +
  geom_boxplot() +
  coord_flip() +
  theme_bw() +
  #ggtitle("The 3 Rules of Real Estate: Location, Location, Location") +
  ylab("Log Price")
```

This motivates treating `Neighborhood` as a random effect.

## Quality scores are mostly linear {.smaller}

\scriptsize
A number of features are either absent or have an ordinal quality score. While there is no scale to the quality scores, their relationship to average price is often linear anyway, so treating them as numerical isn't so far-fetched.

```{r, echo=FALSE}
df_plot <- ames_train %>% select(price, Fireplace.Qu, Exter.Qual, Bsmt.Qual, Garage.Finish, BsmtFin.Type.1, BsmtFin.Type.2)

# Reshape to wide format
dfmelt <- melt(df_plot, measure.vars=setdiff(colnames(df_plot),"price"), value.name = "Score")


ggplot(dfmelt, aes(y=price,x=Score))+
  geom_point(size=.7, alpha=.2) +
  stat_smooth(method = "lm", col = "red") +
  facet_wrap( ~ variable, ncol=2) +
  ylab("Log Price") +
  xlab("Quality Score") +
  #ggtitle("Quality scores are mostly linear") +
  scale_x_continuous(breaks=1:7,label=c("None","Q1\n(Lowest Qual)", "Q2","Q3","Q4","Q5","Q6\n(Highest Qual)"))
```

## PCA

```{r, cache=TRUE, include=FALSE}
modgam = gam(formula = price ~ 
               s(Total.SF,k=20) + 
               Overall.Qual + 
               Total.SF:Overall.Qual+
               s(Year.Built,k=10) + 
               Overall.Cond + 
               s(Bsmt.Unf.SF, k=10) + 
               s(Neighborhood,bs='re') + 
               s(Garage.Area, k=10) + 
               Fireplace.Qu + 
               s(Lot.Area, k=10) + 
               Central.Air + 
               s(Year.Remod.Add, k=10) + 
               Bldg.Type:Central.Air+
               Bldg.Type:Lot.Area+
               Bldg.Type+
               BsmtFin.SF.2 + 
               BsmtFin.SF.1 + 
               Functional+
               Condition.1 + 
               Bsmt.Exposure+
               Full.Bath + 
               Full.Bath:Overall.Qual+
               Paved.Drive +
               Exter.Qual:Overall.Qual+
               Exter.Qual+
               N.Floors+
               Porch:Porch.SF+
               Kitchen.Qual+
               s(Yr.Sold, bs="re"),
             data=ames_train)
```

```{r PCA, echo=FALSE}
numeric = which(sapply(ames_train, is.numeric))
interaction = c("Overall.Qual", "Total.SF", "Lot.Area", "Full.Bath", "Exter.Qual", "Porch.SF")
splines = c("Total.SF", "Year.Built", "Bsmt.Unf.SF", "Garage.Area", "Lot.Area","Year.Remod.Add", "Yr.Sold")
linear = c("Overall.Cond", "Fireplace.Qu", "BsmtFin.SF.2", "BsmtFin.SF.1","Bsmt.Exposure", "Full.Bath","Paved.Drive", "Exter.Qual","N.Floors", "Kitchen.Qual")
pca = prcomp(ames_train[numeric][,-c(1,2)], scale. = TRUE)
#plot(pca, type="l")
pcanames = names(pca$rotation[,1])
type = factor(sapply(pcanames, function(x){
  if(is.element(x,interaction)){return ("Interaction")}
  else if(is.element(x,splines)){return ("Spline")}
  else if(is.element(x,linear)){return ("Linear")}
  else{return ("Not in model")}
}))


pcadf = data.frame(variables = pcanames, y = pca$rotation[,1], type = type)
ggplot(data = pcadf, aes(x=variables, y=y, fill = type))+
  geom_bar(stat="identity")+ coord_flip() + ggtitle("PCA on numerical predictors") +
  ylab("Weight in the first principal component")
```


## GAM


```{r, echo=FALSE}
gamcoef <- coef(modgam)

id_neighborhood <- sapply(names(gamcoef), function(x) grepl("s\\(N",x) )

table.coef.neighborhood <- as.matrix(gamcoef[id_neighborhood])
rownames(table.coef.neighborhood) <- levels(ames_train$Neighborhood)

plot(table.coef.neighborhood, 
     ames_train %>% group_by(Neighborhood) %>% summarize(mean(price)) %>% .[[2]],
     xlab="Neighborhood random effect",
     ylab="Neighborhood average price",
     main="Random effects and average prices by neighborhood")

text(.34, 12.5, "Green Hills", col='red')
```



## Overvalued / Undervalued



```{r, include=FALSE}
# Predictions
predgam.train <- exp(predict(modgam))
predgam.test <- exp(predict(modgam, newdata=ames_test))

# Dataframe of predictions
pred.train <-  data.frame(prediction=predgam.train, 
                          actual=exp(ames_train$price),
                          resid=exp(ames_train$price) - predgam.train,
                          PID = ames_train$PID)

pred.test <-  data.frame(prediction=predgam.test, 
                         actual=exp(ames_test$price),
                         resid=exp(ames_test$price)-predgam.test,
                         PID = ames_test$PID)


```


```{r, include=FALSE}
# Top 10 over/undervalued
top_over <- pred.train %>%
  arrange(desc(resid)) %>%
  slice(1:10) %>%
  left_join(ames_train, by="PID")


top_under <- pred.train %>%
  arrange(resid) %>%
  slice(1:10) %>%
  left_join(ames_train, by="PID")
```




```{r, echo=FALSE, warnings=FALSE, messages=FALSE}
df_over <- top_over %>%
  select(PID, actual, prediction) %>%
  mutate(diff = actual-prediction)

PID_order <- arrange(df_over,desc(diff))$PID

df_over <- melt(df_over, measure.vars = c("actual","prediction"), variable.name = "valuation")

df_over$PID <- factor(df_over$PID, levels = PID_order)

p_over <- ggplot(data = arrange(df_over,valuation), aes(x = as.factor(PID), y = value, fill=valuation)) + 
  geom_bar(stat = "identity", position="dodge")+
  ylab("Actual/Predicted Price")+
  xlab("PID") +
  ggtitle("Top Ten Overvalued Houses")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(legend.justification = c(1,0), legend.position=c(1,0))


df_under <- top_under %>%
  select(PID, actual, prediction) %>%
  mutate(diff = actual-prediction)

PID_order <- arrange(df_under,desc(diff))$PID

df_under <- melt(df_under, measure.vars = c("actual","prediction"), variable.name = "valuation")

df_under$PID <- factor(df_under$PID, levels = PID_order)

p_under <- suppressWarnings(ggplot(data = arrange(df_under,valuation), aes(x = as.factor(PID), y = value, fill=valuation)) + 
  geom_bar(stat = "identity", position="dodge")+
  ylab("")+
  xlab("PID") +
  ggtitle("Top Ten Overvalued Houses")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  theme(legend.justification = c(1,0), legend.position=c(1,0)))

grid.arrange(p_over, p_under, ncol=2)
```






